---
title: "Data_processing"
author: "Changsheng Li"
date: "2024-04-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Available datasets

Currently 9 datases are reviewed and chosen:

| Study code |  GEO ID   | Location | Related study            |
|------------|-----------|----------|--------------------------|
|    1.1     | GSE123902 |   Lung   | Laughney et al. 2020     |
|    1.2     | GSE131907 |   Lung   | Kim et al. 2020          |
|    2.1     | GSE163120 |   Brain  | Pombo Antunes et al 2019 |
|    2.2     | GSE135045 |   Brain  | Xiao et al 2022          |
|    3.1     | GSE118389 |   Breast | Karaayvaz et al 2021     |
|    3.2     | GSE161529 |   Breast | Pal et al 2021           |
|    4.1     | GSE184880 |   Ovary  | Xu et al 2020            |
|    4.2     | GSE154600 |   Ovary  | Geistlinger et al 2021   |


## Data preparation

All datasets would be loaded as they were provided to the database.

The functions used to load the dataset are stored in functions.R.

```{r}
source(file = "functions.R")
library(ggplot2)
```

Use the function `load_seurat(accession, study.code, rawdata.dir, saveddata.dir)` to load dataset individually. 

No default directory to store raw data or Seurat objects. Change the directories in the following chuck.

```{r}
rawdata.dir="~/DFS/RawData/403-Master-RawData/BioInf/Changsheng_Li/Rawdata/"
saveddata.dir="~/DFS/RawData/403-Master-RawData/BioInf/Changsheng_Li/Saveddata/"

fig.path.nFeature.ori = "~/DFS/users/lich/Images/nFeature_original/"
fig.path.nCount.ori = "~/DFS/users/lich/Images/nCount_original/"
fig.path.MT.ori = "~/DFS/users/lich/Images/MT_original/"
```

## Data Processing

### 1. Lung

#### 1.1 GSE123902

```{r}
object <- load_seurat(study.code = 1.1,
                      rawdata.dir = rawdata.dir,
                      saveddata.dir = saveddata.dir)
```

```{r fig.height=10, fig.width=10}
object[["percent.mt"]] <- PercentageFeatureSet(object, pattern = "^MT-")

p1 = VlnPlot(object, features = "nFeature_RNA",alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nFeature - 1.1") +  
  geom_hline(yintercept = 200, linetype="dashed", color = "red") +
  geom_hline(yintercept = 5000, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,10000, length.out=5), 200, 5000)))+
  coord_cartesian(ylim = c(0, 10000))
p1
ggsave("1.1_nFeature.eps",plot = p1, device = "eps", path = fig.path.nFeature.ori)

p2 = VlnPlot(object, features = "nCount_RNA", alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nCount_RNA - 1.1")
p2

ggsave("1.1_nCount.eps",plot = p2, device = "eps", path = fig.path.nCount.ori)

p3 = VlnPlot(object, features = "percent.mt", alpha = 0) +
  theme(legend.position = 'none') +
  ggtitle("MT% - 1.1") +
  coord_cartesian(ylim = c(0, 50)) +
  geom_hline(yintercept = 10, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,50, length.out=6), 10)))
p3
ggsave("1.1_MT.eps",plot = p3, device = "eps", path = fig.path.MT.ori)

```



```{r fig.width=10}

plot1 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

plot1 + plot2
```

```{r}

object_subset <- subset(object, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mt < 10 & tissue == "Patient primary lung adenocarcinoma")
VlnPlot(object_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), combine = F)
```


##### Nomalization

```{r fig.width=10}
object_subset <- NormalizeData(object_subset)

object_subset <- FindVariableFeatures(object_subset, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(object_subset), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(object_subset)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```


##### Scaling

```{r}
all.genes <- rownames(object_subset)
object_subset <- ScaleData(object_subset, features = all.genes)
```

##### DR

PCA
```{r}
object_subset <- RunPCA(object_subset, features = VariableFeatures(object = object_subset))
VizDimLoadings(object_subset, dims = 1:2, reduction = "pca")
DimPlot(object_subset, reduction = "pca")
ElbowPlot(object_subset)
```

UMAP

```{r}
object_subset <- RunUMAP(object_subset, dims = 1:20)
DimPlot(object_subset, reduction = "umap", alpha=1)
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_UMAP/1.1_UMAP.RDS"))
```

##### Plot features

tumor related genes

```{r}
FeaturePlot(object = object_subset, features = c("EPCAM","KRT19","CEA", "SCC","AIF1"))
```

##### data integration

```{r}
object_subset <- readRDS(file = paste0(saveddata.dir, "Seurat_object_UMAP/1.1_UMAP.RDS"))
#object_subset[["RNA"]] <- split(object_subset[["RNA"]], f=object_subset$patient)

object_subset <- SCTransform(object_subset, vst.flavor = "v2")
object_subset <- RunPCA(object_subset, verbose = FALSE)

object_subset <- IntegrateLayers(object = object_subset, method = CCAIntegration, orig.reduction = "pca", new.reduction = "integrated.cca", verbose = FALSE)

# re-join layers after integration
object_subset[["RNA"]] <- JoinLayers(object_subset[["RNA"]])

object_subset <- FindNeighbors(object_subset, reduction = "integrated.cca", dims = 1:30)
object_subset <- FindClusters(object_subset, resolution = 0.1)
object_subset <- RunUMAP(object_subset, dims = 1:50, reduction = "integrated.cca")
DimPlot(object_subset, reduction = "umap", group.by = c("geo_accession", "seurat_clusters"))
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_integrated_UMAP/1.1_integrated_UMAP.RDS"))
```

#### 1.2 GSE131907

```{r}
object <- load_seurat(study.code = 1.2,
                      rawdata.dir = rawdata.dir,
                      saveddata.dir = saveddata.dir)
```

```{r fig.height=10, fig.width=10}
Idents(object) <- "Sample"
object[["percent.mt"]] <- PercentageFeatureSet(object, pattern = "^MT-")

p1 = VlnPlot(object, features = "nFeature_RNA",alpha = 0,layer = "counts") + 
  theme(legend.position = 'none') +
  ggtitle("nFeature - 1.2") +  
  geom_hline(yintercept = 200, linetype="dashed", color = "red") +
  geom_hline(yintercept = 5000, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,10000, length.out=5), 200, 5000)))+
  coord_cartesian(ylim = c(0, 10000))
p1
ggsave("1.2_nFeature.eps",plot = p1, device = "eps", path = fig.path.nFeature.ori)

p2 = VlnPlot(object, features = "nCount_RNA", alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nCount_RNA - 1.2")
p2
ggsave("1.2_nCount.eps",plot = p2, device = "eps", path = fig.path.nCount.ori)

p3 = VlnPlot(object, features = "percent.mt", alpha = 0) +
  theme(legend.position = 'none') +
  ggtitle("MT% - 1.2") +
  coord_cartesian(ylim = c(0, 50)) +
  geom_hline(yintercept = 10, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,50, length.out=6), 10)))
p3
ggsave("1.2_MT.eps",plot = p3, device = "eps", path = fig.path.MT.ori)

```



```{r fig.width=10}

plot1 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

plot1 + plot2

```

```{r}
object_subset <- subset(object, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mt < 10 & Sample_Origin == c("tL/B", "tLung"))
VlnPlot(object_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), combine = F, alpha = 0)
```

##### Normalization

```{r fig.width=10}
object_subset <- split(object_subset, f = object_subset$Sample)
object_subset <- NormalizeData(object_subset,normalization.method = "LogNormalize" )

object_subset <- FindVariableFeatures(object_subset, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(object_subset), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(object_subset)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```


##### Scaling

```{r}
all.genes <- rownames(object_subset)
object_subset <- ScaleData(object_subset, features = all.genes, do.scale = F, do.center = T, scale.max=10)
```

##### DR

PCA
```{r}
object_subset <- RunPCA(object_subset, features = VariableFeatures(object = object_subset))
VizDimLoadings(object_subset, dims = 1:2, reduction = "pca")
DimPlot(object_subset, reduction = "pca", dims = 1:2, group.by = "Cell_type")
ElbowPlot(object_subset)
```

UMAP

```{r}
object_subset <- RunUMAP(object_subset, dims = 1:30)
undet <- WhichCells(object = object_subset,  c(object_subset$Cell_type == c("Epithelial cells")))
DimPlot(object_subset, reduction = "umap", cells.highlight = c(cells.highlight = undet))
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_UMAP/1.2_UMAP.RDS"))
```

##### Plot features

tumor related genes

```{r}
FeaturePlot(object = object_subset, features = c("EPCAM","KRT19","CEA", "SCC","AIF1"))
```

##### data integration

```{r}
object_subset <- readRDS(file = paste0(saveddata.dir, "Seurat_object_UMAP/1.2_UMAP.RDS"))
object_subset[["RNA"]] <- split(object_subset[["RNA"]], f=object_subset$Sample)

#object_subset <- SCTransform(object_subset, vst.flavor = "v2")
#object_subset <- RunPCA(object_subset, verbose = FALSE)

object_subset <- IntegrateLayers(object = object_subset, method = CCAIntegration, orig.reduction = "pca", new.reduction = "integrated.cca", verbose = FALSE)

# re-join layers after integration
object_subset[["RNA"]] <- JoinLayers(object_subset[["RNA"]])

object_subset <- FindNeighbors(object_subset, reduction = "integrated.cca", dims = 1:30)
object_subset <- FindClusters(object_subset, resolution = 0.01)
object_subset <- RunUMAP(object_subset, dims = 1:30, reduction = "integrated.cca")
DimPlot(object_subset, reduction = "umap", group.by = c("geo_accession", "seurat_clusters"))
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_integrated_UMAP/1.2_integrated_UMAP.RDS"))
```


### 2. Brain

#### 2.1 GSE163120 (we only take the subsample GSM4972211[human])

```{r}
object <- load_seurat(study.code = 2.1,
                      rawdata.dir = rawdata.dir,
                      saveddata.dir = saveddata.dir)
```

```{r fig.height=10, fig.width=10}
object[["percent.mt"]] <- PercentageFeatureSet(object, pattern = "^MT-")

p1 = VlnPlot(object, features = "nFeature_RNA",alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nFeature - 2.1") +  
  geom_hline(yintercept = 200, linetype="dashed", color = "red") +
  geom_hline(yintercept = 5000, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,10000, length.out=5), 200, 5000)))+
  coord_cartesian(ylim = c(0, 10000))
p1
ggsave("2.1_nFeature.eps",plot = p1, device = "eps", path = fig.path.nFeature.ori)

p2 = VlnPlot(object, features = "nCount_RNA", alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nCount_RNA - 2.1")
p2
ggsave("2.1_nCount.eps",plot = p2, device = "eps", path = fig.path.nCount.ori)

p3 = VlnPlot(object, features = "percent.mt", alpha = 0) +
  theme(legend.position = 'none') +
  ggtitle("MT% - 2.1") +
  coord_cartesian(ylim = c(0, 50)) +
  geom_hline(yintercept = 10, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,50, length.out=6), 10)))
p3
ggsave("2.1_MT.eps",plot = p3, device = "eps", path = fig.path.MT.ori)

```


```{r fig.width=10}

plot1 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

plot1 + plot2

```
```{r}
object_subset <- subset(object, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mt < 10)
VlnPlot(object_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), combine = F)
```

##### Normalization

```{r fig.width=10}
object_subset <- NormalizeData(object_subset)

object_subset <- FindVariableFeatures(object_subset, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(object_subset), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(object_subset)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```


##### Scaling

```{r}
all.genes <- rownames(object_subset)
object_subset <- ScaleData(object_subset, features = all.genes)
```

##### DR

PCA
```{r}
object_subset <- RunPCA(object_subset, features = VariableFeatures(object = object_subset))
VizDimLoadings(object_subset, dims = 1:2, reduction = "pca")
DimPlot(object_subset, reduction = "pca")
ElbowPlot(object_subset)
```

UMAP

```{r}
object_subset <- RunUMAP(object_subset, dims = 1:20)
DimPlot(object_subset, reduction = "umap", group.by = "cluster")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_UMAP/2.1_UMAP.RDS"))
```

##### Plot features

tumor related genes

```{r}
FeaturePlot(object = object_subset, features = c("EPCAM","KRT19","CEA", "SCC","AIF1"))
```

##### data integration

```{r}
object_subset <- readRDS(file = paste0(saveddata.dir, "Seurat_object_UMAP/2.1_UMAP.RDS"))
object_subset[["RNA"]] <- split(object_subset[["RNA"]], f=object_subset$sample)

object_subset <- SCTransform(object_subset, vst.flavor = "v2")
object_subset <- RunPCA(object_subset, verbose = FALSE)

object_subset <- IntegrateLayers(object = object_subset, method = CCAIntegration, orig.reduction = "pca", new.reduction = "integrated.cca", verbose = FALSE)

# re-join layers after integration
object_subset[["RNA"]] <- JoinLayers(object_subset[["RNA"]])

object_subset <- FindNeighbors(object_subset, reduction = "integrated.cca", dims = 1:30)
object_subset <- FindClusters(object_subset, resolution = 0.2)
object_subset <- RunUMAP(object_subset, dims = 1:30, reduction = "integrated.cca")
DimPlot(object_subset, reduction = "umap", group.by = c("geo_accession", "seurat_clusters"))
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_integrated_UMAP/2.1_integrated_UMAP.RDS"))
```


#### 2.2 GSE182109

```{r}
object <- load_seurat(study.code = 2.2,
                      rawdata.dir = rawdata.dir,
                      saveddata.dir = saveddata.dir)
```

```{r fig.height=10, fig.width=10}
object[["percent.mt"]] <- PercentageFeatureSet(object, pattern = "^MT-")

p1 = VlnPlot(object, features = "nFeature_RNA",alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nFeature - 2.2") +  
  geom_hline(yintercept = 200, linetype="dashed", color = "red") +
  geom_hline(yintercept = 5000, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,10000, length.out=5), 200, 5000)))+
  coord_cartesian(ylim = c(0, 10000))
p1
ggsave("2.2_nFeature.eps",plot = p1, device = "eps", path = fig.path.nFeature.ori)

p2 = VlnPlot(object, features = "nCount_RNA", alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nCount_RNA - 2.2")
p2
ggsave("2.2_nCount.eps",plot = p2, device = "eps", path = fig.path.nCount.ori)

p3 = VlnPlot(object, features = "percent.mt", alpha = 0) +
  theme(legend.position = 'none') +
  ggtitle("MT% - 2.2") +
  coord_cartesian(ylim = c(0, 50)) +
  geom_hline(yintercept = 10, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,50, length.out=6), 10)))
p3
ggsave("2.2_MT.eps",plot = p3, device = "eps", path = fig.path.MT.ori)

```

```{r fig.width=10}

plot1 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

plot1 + plot2

```
```{r}
object_subset <- subset(object, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mt < 10)
VlnPlot(object_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), combine = F, alpha = 0)
```

##### Normalization

```{r fig.width=10}
object_subset <- NormalizeData(object_subset)

object_subset <- FindVariableFeatures(object_subset, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(object_subset), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(object_subset)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```


##### Scaling

```{r}
all.genes <- rownames(object_subset)
object_subset <- ScaleData(object_subset, features = all.genes)
```

##### DR

PCA
```{r}
object_subset <- RunPCA(object_subset, features = VariableFeatures(object = object_subset))
VizDimLoadings(object_subset, dims = 1:2, reduction = "pca")
DimPlot(object_subset, reduction = "pca")
ElbowPlot(object_subset)
```

UMAP

```{r}
object_subset <- RunUMAP(object_subset, dims = 1:20)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_UMAP/2.2_UMAP.RDS"))
```

##### Plot features

tumor related genes

```{r}
FeaturePlot(object = object_subset, features = c("EPCAM","KRT19","P2RY12", "GRN","AIF1"))
```

##### data integration (no need)

```{r}
object_subset <- readRDS(file = paste0(saveddata.dir, "Seurat_object_UMAP/2.2_UMAP.RDS"))
#object_subset[["RNA"]] <- split(object_subset[["RNA"]], f=object_subset$sample)

object_subset <- FindNeighbors(object_subset, reduction = "pca", dims = 1:50)
object_subset <- FindClusters(object_subset, resolution = 0.2)
object_subset <- RunUMAP(object_subset, dims = 1:50, reduction = "pca")
DimPlot(object_subset, reduction = "umap", group.by = c("geo_accession", "seurat_clusters"))
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_integrated_UMAP/2.2_integrated_UMAP.RDS"))
```

### 3. Breast

#### 3.1 GSE75688

```{r}
object <- load_seurat(study.code = 3.1,
                      rawdata.dir = rawdata.dir,
                      saveddata.dir = saveddata.dir)
```

```{r fig.height=10, fig.width=10}
Idents(object) <- "patient"
object[["percent.mt"]] <- PercentageFeatureSet(object, pattern = "^MT-")

p1 = VlnPlot(object, features = "nFeature_RNA",alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nFeature - 3.1") +  
  geom_hline(yintercept = 200, linetype="dashed", color = "red") +
  geom_hline(yintercept = 5000, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,10000, length.out=5), 200, 5000)))+
  coord_cartesian(ylim = c(0, 10000))
p1
ggsave("3.1_nFeature.eps",plot = p1, device = "eps", path = fig.path.nFeature.ori)

p2 = VlnPlot(object, features = "nCount_RNA", alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nCount_RNA - 3.1")
p2
ggsave("3.1_nCount.eps",plot = p2, device = "eps", path = fig.path.nCount.ori)

p3 = VlnPlot(object, features = "percent.mt", alpha = 0) +
  theme(legend.position = 'none') +
  ggtitle("MT% - 3.1") +
  coord_cartesian(ylim = c(0, 50)) +
  geom_hline(yintercept = 10, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,50, length.out=6), 10)))
p3
ggsave("3.1_MT.eps",plot = p3, device = "eps", path = fig.path.MT.ori)

```


```{r fig.width=10}

plot1 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

plot1 + plot2

```
```{r}
object_subset <- subset(object, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mt < 10)
VlnPlot(object_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), combine = F)
```

##### Normalization

```{r fig.width=10}
object_subset <- NormalizeData(object_subset)

object_subset <- FindVariableFeatures(object_subset, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(object_subset), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(object_subset)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```


##### Scaling

```{r}
all.genes <- rownames(object_subset)
object_subset <- ScaleData(object_subset, features = all.genes)
```

##### DR

PCA
```{r}
object_subset <- RunPCA(object_subset, features = VariableFeatures(object = object_subset))
VizDimLoadings(object_subset, dims = 1:2, reduction = "pca")
DimPlot(object_subset, reduction = "pca")
ElbowPlot(object_subset)
```

UMAP

```{r}
object_subset <- RunUMAP(object_subset, dims = 1:20)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_UMAP/3.1_UMAP.RDS"))
```

##### Plot features

tumor related genes

```{r}
object_subset <- readRDS(file = paste0(saveddata.dir, "Seurat_object_UMAP/3.1_UMAP.RDS"))
FeaturePlot(object = object_subset, features = c("EPCAM","KRT19","CEA", "SCC","AIF1"))
```

##### data integration

```{r}
object_subset[["RNA"]] <- split(object_subset[["RNA"]], f=object_subset$orig.ident)

object_subset <- SCTransform(object_subset, vst.flavor = "v2")
object_subset <- RunPCA(object_subset, verbose = FALSE)

object_subset <- IntegrateLayers(object = object_subset, method = CCAIntegration, orig.reduction = "pca", new.reduction = "integrated.cca", verbose = FALSE, k.weight = 50)

# re-join layers after integration
object_subset[["RNA"]] <- JoinLayers(object_subset[["RNA"]])

object_subset <- FindNeighbors(object_subset, reduction = "integrated.cca", dims = 1:30)
object_subset <- FindClusters(object_subset, resolution = 1)
object_subset <- RunUMAP(object_subset, dims = 1:30, reduction = "integrated.cca")
DimPlot(object_subset, reduction = "umap", group.by = "orig.ident")

saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_integrated_UMAP/3.1_integrated_UMAP.RDS"))
```



#### 3.2 GSE161529

```{r}
object <- load_seurat(study.code = 3.2,
                      rawdata.dir = rawdata.dir,
                      saveddata.dir = saveddata.dir)
```

```{r fig.height=10, fig.width=10}
Idents(object) <- "patient"
object[["percent.mt"]] <- PercentageFeatureSet(object, pattern = "^MT-")

p1 = VlnPlot(object, features = "nFeature_RNA",alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nFeature - 3.2") +  
  geom_hline(yintercept = 200, linetype="dashed", color = "red") +
  geom_hline(yintercept = 5000, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,10000, length.out=5), 200, 5000)))+
  coord_cartesian(ylim = c(0, 10000))
p1
ggsave("3.2_nFeature.eps",plot = p1, device = "eps", path = fig.path.nFeature.ori)

p2 = VlnPlot(object, features = "nCount_RNA", alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nCount_RNA - 3.2")
p2
ggsave("3.2_nCount.eps",plot = p2, device = "eps", path = fig.path.nCount.ori)

p3 = VlnPlot(object, features = "percent.mt", alpha = 0) +
  theme(legend.position = 'none') +
  ggtitle("MT% - 3.2") +
  coord_cartesian(ylim = c(0, 50)) +
  geom_hline(yintercept = 10, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,50, length.out=6), 10)))
p3
ggsave("3.2_MT.eps",plot = p3, device = "eps", path = fig.path.MT.ori)

```

```{r fig.width=10}

plot1 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

plot1 + plot2

```
```{r}
object_subset <- subset(object, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mt < 10 & cancer.type != "Normal" & cell.population == "Total")
VlnPlot(object_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), combine = F, alpha = 0)
```

##### Normalization

```{r fig.width=10}
object_subset <- NormalizeData(object_subset)

object_subset <- FindVariableFeatures(object_subset, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(object_subset), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(object_subset)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```


##### Scaling

```{r}
all.genes <- rownames(object_subset)
object_subset <- ScaleData(object_subset, features = all.genes)
```

##### DR

PCA
```{r}
object_subset <- RunPCA(object_subset, features = VariableFeatures(object = object_subset))
VizDimLoadings(object_subset, dims = 1:2, reduction = "pca")
DimPlot(object_subset, reduction = "pca")
ElbowPlot(object_subset)
```

UMAP

```{r}
object_subset <- RunUMAP(object_subset, dims = 1:20)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_UMAP/3.2_UMAP.RDS"))
```
##### Plot features

tumor related genes

```{r}
FeaturePlot(object = object_subset, features = c("EPCAM","KRT19","CEA", "SCC","AIF1"))
```

##### data integration

```{r}
object_subset <- readRDS(file = paste0(saveddata.dir, "Seurat_object_UMAP/3.2_UMAP.RDS"))
#object_subset[["RNA"]] <- split(object_subset[["RNA"]], f=object_subset$sample)

object_subset <- SCTransform(object_subset, vst.flavor = "v2")
object_subset <- RunPCA(object_subset, verbose = FALSE)

object_subset <- IntegrateLayers(object = object_subset, method = CCAIntegration, orig.reduction = "pca", new.reduction = "integrated.cca", verbose = FALSE)

# re-join layers after integration
object_subset[["RNA"]] <- JoinLayers(object_subset[["RNA"]])

object_subset <- FindNeighbors(object_subset, reduction = "integrated.cca", dims = 1:30)
object_subset <- FindClusters(object_subset, resolution = 0.2)
object_subset <- RunUMAP(object_subset, dims = 1:30, reduction = "integrated.cca")
DimPlot(object_subset, reduction = "umap", group.by = c("seurat_clusters"))
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_integrated_UMAP/3.2_integrated_UMAP.RDS"))
```


### 4. Ovary

#### 4.1 GSE184880

```{r}
object <- load_seurat(study.code = 4.1,
                      rawdata.dir = rawdata.dir,
                      saveddata.dir = saveddata.dir)
```

```{r fig.height=10, fig.width=10}
Idents(object) <- "geo_accession"
object[["percent.mt"]] <- PercentageFeatureSet(object, pattern = "^MT-")

p1 = VlnPlot(object, features = "nFeature_RNA",alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nFeature - 4.1") +  
  geom_hline(yintercept = 200, linetype="dashed", color = "red") +
  geom_hline(yintercept = 5000, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,10000, length.out=5), 200, 5000)))+
  coord_cartesian(ylim = c(0, 10000))
p1
ggsave("4.1_nFeature.eps",plot = p1, device = "eps", path = fig.path.nFeature.ori)

p2 = VlnPlot(object, features = "nCount_RNA", alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nCount_RNA - 4.1")
p2
ggsave("4.1_nCount.eps",plot = p2, device = "eps", path = fig.path.nCount.ori)

p3 = VlnPlot(object, features = "percent.mt", alpha = 0) +
  theme(legend.position = 'none') +
  ggtitle("MT% - 4.1") +
  coord_cartesian(ylim = c(0, 50)) +
  geom_hline(yintercept = 10, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,50, length.out=6), 10)))
p3
ggsave("4.1_MT.eps",plot = p3, device = "eps", path = fig.path.MT.ori)

```


```{r fig.width=10}

plot1 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

plot1 + plot2

```
```{r}
object_subset <- subset(object, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mt < 10)
VlnPlot(object_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), combine = F)
```

##### Normalization

```{r fig.width=10}
object_subset <- NormalizeData(object_subset)

object_subset <- FindVariableFeatures(object_subset, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(object_subset), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(object_subset)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```



##### Scaling

```{r}
all.genes <- rownames(object_subset)
object_subset <- ScaleData(object_subset, features = all.genes)
```

##### DR

PCA
```{r}
object_subset <- RunPCA(object_subset, features = VariableFeatures(object = object_subset))
VizDimLoadings(object_subset, dims = 1:2, reduction = "pca")
DimPlot(object_subset, reduction = "pca")
ElbowPlot(object_subset, ndims = 30)
```

UMAP

```{r}
object_subset <- RunUMAP(object_subset, dims = 1:30)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_UMAP/4.1_UMAP.RDS"))
```
##### Plot features

tumor related genes

```{r}
FeaturePlot(object = object_subset, features = c("EPCAM","KRT19","CEA", "SCC","AIF1"))
```

##### data integration

```{r}
object_subset <- readRDS(file = paste0(saveddata.dir, "Seurat_object_UMAP/4.1_UMAP.RDS"))
#object_subset[["RNA"]] <- split(object_subset[["RNA"]], f=object_subset$patient)

object_subset <- SCTransform(object_subset, vst.flavor = "v2")
object_subset <- RunPCA(object_subset, verbose = FALSE)

object_subset <- IntegrateLayers(object = object_subset, method = CCAIntegration, orig.reduction = "pca", new.reduction = "integrated.cca", verbose = FALSE)

# re-join layers after integration
object_subset[["RNA"]] <- JoinLayers(object_subset[["RNA"]])

object_subset <- FindNeighbors(object_subset, reduction = "integrated.cca", dims = 1:30)
object_subset <- FindClusters(object_subset, resolution = 0.01)
object_subset <- RunUMAP(object_subset, dims = 1:30, reduction = "integrated.cca")
DimPlot(object_subset, reduction = "umap", group.by = c("geo_accession", "seurat_clusters"))
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_integrated_UMAP/4.1_integrated_UMAP.RDS"))
```


#### 4.2 GSE154600

```{r}
object <- load_seurat(study.code = 4.2,
                      rawdata.dir = rawdata.dir,
                      saveddata.dir = saveddata.dir)
```

```{r fig.height=10, fig.width=10}
Idents(object) <- "geo_accession"
object[["percent.mt"]] <- PercentageFeatureSet(object, pattern = "^MT-")

p1 = VlnPlot(object, features = "nFeature_RNA",alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nFeature - 4.2") +  
  geom_hline(yintercept = 200, linetype="dashed", color = "red") +
  geom_hline(yintercept = 5000, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,10000, length.out=5), 200, 5000)))+
  coord_cartesian(ylim = c(0, 10000))
p1
ggsave("4.2_nFeature.eps",plot = p1, device = "eps", path = fig.path.nFeature.ori)

p2 = VlnPlot(object, features = "nCount_RNA", alpha = 0) + 
  theme(legend.position = 'none') +
  ggtitle("nCount_RNA - 4.2")
p2
ggsave("4.2_nCount.eps",plot = p2, device = "eps", path = fig.path.nCount.ori)

p3 = VlnPlot(object, features = "percent.mt", alpha = 0) +
  theme(legend.position = 'none') +
  ggtitle("MT% - 4.2") +
  coord_cartesian(ylim = c(0, 50)) +
  geom_hline(yintercept = 10, linetype="dashed", color = "red") +
  scale_y_continuous(breaks = sort(c(seq(0,50, length.out=6), 10)))
p3
ggsave("4.2_MT.eps",plot = p3, device = "eps", path = fig.path.MT.ori)

```


```{r fig.width=10}

plot1 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

plot1 + plot2

```

```{r}
object_subset <- subset(object, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mt < 10)
VlnPlot(object_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), combine = F)
```

##### Normalization

```{r fig.width=10}
object_subset <- NormalizeData(object_subset)

object_subset <- FindVariableFeatures(object_subset, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(object_subset), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(object_subset)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```


##### Scaling

```{r}
all.genes <- rownames(object_subset)
object_subset <- ScaleData(object_subset, features = all.genes)
```

##### DR

PCA
```{r}
object_subset <- RunPCA(object_subset, features = VariableFeatures(object = object_subset))
VizDimLoadings(object_subset, dims = 1:2, reduction = "pca")
DimPlot(object_subset, reduction = "pca")
ElbowPlot(object_subset)
```

UMAP

```{r}
object_subset <- RunUMAP(object_subset, dims = 1:20)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_UMAP/4.2_UMAP.RDS"))
```
##### Plot features

tumor related genes

```{r}
FeaturePlot(object = object_subset, features = c("EPCAM","KRT19","CEA", "SCC","AIF1"))
```

##### data integration

```{r}
object_subset <- readRDS(file = paste0(saveddata.dir, "Seurat_object_UMAP/4.2_UMAP.RDS"))
#object_subset[["RNA"]] <- split(object_subset[["RNA"]], f=object_subset$patient)

object_subset <- SCTransform(object_subset, vst.flavor = "v2")
object_subset <- RunPCA(object_subset, verbose = FALSE)

object_subset <- IntegrateLayers(object = object_subset, method = CCAIntegration, orig.reduction = "pca", new.reduction = "integrated.cca", verbose = FALSE)

# re-join layers after integration
object_subset[["RNA"]] <- JoinLayers(object_subset[["RNA"]])

object_subset <- FindNeighbors(object_subset, reduction = "integrated.cca", dims = 1:30)
object_subset <- FindClusters(object_subset, resolution = 0.1)
object_subset <- RunUMAP(object_subset, dims = 1:50, reduction = "integrated.cca")
DimPlot(object_subset, reduction = "umap", group.by = c("geo_accession", "seurat_clusters"))
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_integrated_UMAP/4.2_integrated_UMAP.RDS"))
```