---
title: "Macrophages_selection"
author: "Changsheng Li"
date: "2024-06-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation

All datasets would be loaded with UMAP.

The functions used to load the dataset are stored in functions.R.

```{r}
source(file = "functions.R")
library(ggplot2)
```

Use the function `load_seurat(accession, after_UMAP = T, study.code, rawdata.dir, saveddata.dir)` to load integrated data sets with UMAP individually. 

No default directory to store raw data or Seurat objects. Change the directories in the following chuck.

```{r}
saveddata.dir="~/DFS/RawData/403-Master-RawData/BioInf/Changsheng_Li/Saveddata/"
```

## Lung

### 1.1 GSE123902

#### Load dataset
```{r}
object <- load_seurat(study.code = 1.1, after_UMAP = T, saveddata.dir = saveddata.dir)
```

#### UMAP Visualization
```{r fig.height=4,fig.width=5}
DimPlot(object, reduction = "umap")
```

```{r}
#object <- FindNeighbors(object, dims = 1:30)
object <- FindClusters(object, resolution = 0.1)
```

```{r fig.width=12, fig.height=4}
p1 <- DimPlot(object, reduction = "umap", group.by = "geo_accession", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank()) + ggtitle("",subtitle = "GSE123902 (lung 1)")

p2 <- DimPlot(object, reduction = "umap", group.by = "seurat_clusters", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())

p3 <- FeaturePlot(object = object, features = c("AIF1"), alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())
p1 + p2 + p3

ggsave("1.1_cell_filtering_UMAP.pdf",plot = p1+p2+p3, device = "pdf", path = paste0(fig.path, "cell_filtering/"), height = 4, width = 12)

```


#### filter Myeloids

```{r}
object_subset <- subset(object, idents = 2)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_myeloids_integrated/1.1_myeloids_integrated.RDS"))
```


### 1.2 GSE131907


#### Load dataset
```{r}
object <- load_seurat(study.code = 1.2, after_UMAP = T, saveddata.dir = saveddata.dir)
```

#### UMAP Visualization
```{r fig.height=4,fig.width=5}
DimPlot(object, reduction = "umap")
```

```{r}
object <- FindNeighbors(object, dims = 1:30)
object <- FindClusters(object, resolution = 0.05)
```

```{r fig.width=12, fig.height=4}
p1 <- DimPlot(object, reduction = "umap", group.by = "Sample", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank()) + ggtitle("",subtitle = "GSE131907 (lung 2)")

p2 <- DimPlot(object, reduction = "umap", group.by = "seurat_clusters", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())

p3 <- FeaturePlot(object = object, features = c("AIF1"), alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())
p1 + p2 + p3

ggsave("1.2_cell_filtering_UMAP.pdf",plot = p1+p2+p3, device = "pdf", path = paste0(fig.path, "cell_filtering/"), height = 4, width = 12)

```


#### filter Myeloids

```{r}
object_subset <- subset(object, idents = 1)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_myeloids_integrated/1.2_myeloids_integrated.RDS"))
```



## Brain

### 2.1 GSE163120 (GSM4972211)

#### Load dataset
```{r}
object <- load_seurat(study.code = 2.1, after_UMAP = T, saveddata.dir = saveddata.dir)
```

#### UMAP Visualization
```{r fig.height=4,fig.width=5}
DimPlot(object, reduction = "umap")
```

```{r}
object <- FindNeighbors(object, dims = 1:30)
object <- FindClusters(object, resolution = 0.1)
```

```{r fig.width=12, fig.height=4}
p1 <- DimPlot(object, reduction = "umap", group.by = "sample", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank()) + ggtitle("",subtitle = "GSM4972211 (brain 1)")

p2 <- DimPlot(object, reduction = "umap", group.by = "seurat_clusters", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())

p3 <- FeaturePlot(object = object, features = c("AIF1"), alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())


p1 + p2 + p3

ggsave("2.1_cell_filtering_UMAP.pdf",plot = p1+p2+p3, device = "pdf", path = paste0(fig.path, "cell_filtering/"), height = 4, width = 12)

p4 <- DimPlot(object, reduction = "umap", group.by = "cluster", alpha = 0.5) + ggtitle("cell types")

ggsave("2.1_cell_filtering_UMAP_identity.pdf",plot = p4, device = "pdf", path = paste0(fig.path, "cell_filtering/"), height = 4, width = 5)
```


#### filter Myeloids

```{r}
object_subset <- subset(object, idents = c(0, 1, 3))
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_myeloids_integrated/2.1_myeloids_integrated.RDS"))
```



### 2.2 GSE135045 (try to integrate the data to check th eproblem of ndGBM-10)

#### Load dataset
```{r}
object <- load_seurat(study.code = 2.2, after_UMAP = T, saveddata.dir = saveddata.dir)
```

#### UMAP Visualization
```{r fig.height=4,fig.width=5}
DimPlot(object, reduction = "umap")
```

```{r}
object <- FindNeighbors(object, dims = 1:30)
object <- FindClusters(object, resolution = 0.05)

```

```{r fig.width=12, fig.height=4}
p1 <- DimPlot(object, reduction = "umap", group.by = "geo_accession", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank()) + ggtitle("",subtitle = "GSE135045 (brain 2)")

p2 <- DimPlot(object, reduction = "umap", group.by = "seurat_clusters", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())

p3 <- FeaturePlot(object = object, features = c("AIF1"), alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())
p1 + p2 + p3

ggsave("2.2_cell_filtering_UMAP.pdf",plot = p1+p2+p3, device = "pdf", path = paste0(fig.path, "cell_filtering/"), height = 4, width = 12)

```


#### filter Myeloids

```{r}
object_subset <- subset(object, idents = 3)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_myeloids_integrated/2.2_myeloids_integrated.RDS"))
```


## Breast

### 3.1 GSE118389

(Try integration prior to the clustering. If it works, re-run the other datasets and do te same integation)

#### Load dataset
```{r}
object <- load_seurat(study.code = 3.1, after_UMAP = T, saveddata.dir = saveddata.dir)
```

#### UMAP Visualization
```{r fig.height=4,fig.width=5}
DimPlot(object, reduction = "umap")
```

```{r}
object <- FindNeighbors(object, dims = 1:30)
object <- FindClusters(object, resolution = 0.1)
```

```{r fig.width=12, fig.height=4}
p1 <- DimPlot(object, reduction = "umap", group.by = "geo_accession", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank()) + ggtitle("",subtitle = "GSE118389 (breast 1)")

p2 <- DimPlot(object, reduction = "umap", group.by = "seurat_clusters", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())

p3 <- FeaturePlot(object = object, features = c("AIF1"), alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())
p1 + p2 + p3

ggsave("3.1_cell_filtering_UMAP.pdf",plot = p1+p2+p3, device = "pdf", path = paste0(fig.path, "cell_filtering/"), height = 4, width = 12)

```


#### filter Myeloids

```{r}
object_subset <- subset(object, idents = 6)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_myeloids_integrated/3.1_myeloids_integrated.RDS"))
```


### 3.2 GSE161529

#### Load dataset
```{r}
object <- load_seurat(study.code = 3.2, after_UMAP = T, saveddata.dir = saveddata.dir)
```

#### UMAP Visualization
```{r fig.height=4,fig.width=5}
DimPlot(object, reduction = "umap")
```

```{r}
object <- FindNeighbors(object, dims = 1:30)
object <- FindClusters(object, resolution = 0.1)
```

```{r fig.width=12, fig.height=4}
p1 <- DimPlot(object, reduction = "umap", group.by = "geo_accession", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank()) + ggtitle("",subtitle = "GSE161529 (breast 2)")

p2 <- DimPlot(object, reduction = "umap", group.by = "seurat_clusters", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())

p3 <- FeaturePlot(object = object, features = c("AIF1"), alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())
p1 + p2 + p3

ggsave("3.2_cell_filtering_UMAP.pdf",plot = p1+p2+p3, device = "pdf", path = paste0(fig.path, "cell_filtering/"), height = 4, width = 12)

```


#### filter Myeloids

```{r}
object_subset <- subset(object, idents = 2)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_myeloids_integrated/3.2_myeloids_integrated.RDS"))
```



## Ovary

### 4.1 GSE184880

#### Load dataset
```{r}
object <- load_seurat(study.code = 4.1, after_UMAP = T, saveddata.dir = saveddata.dir)
```

#### UMAP Visualization
```{r fig.height=4,fig.width=5}
DimPlot(object, reduction = "umap")
```

```{r}
object <- FindNeighbors(object, dims = 1:30)
object <- FindClusters(object, resolution = 0.1)
```

```{r fig.width=12, fig.height=4}
p1 <- DimPlot(object, reduction = "umap", group.by = "geo_accession", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank()) + ggtitle("",subtitle = "GSE184880 (ovary 1)")

p2 <- DimPlot(object, reduction = "umap", group.by = "seurat_clusters", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())

p3 <- FeaturePlot(object = object, features = c("AIF1"), alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())
p1 + p2 + p3

ggsave("4.1_cell_filtering_UMAP.pdf",plot = p1+p2+p3, device = "pdf", path = paste0(fig.path, "cell_filtering/"), height = 4, width = 12)

```


#### filter Myeloids

```{r}
object_subset <- subset(object, idents = 2)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_myeloids_integrated/4.1_myeloids_integrated.RDS"))
```



### 4.2 GSE154600

#### Load dataset
```{r}
object <- load_seurat(study.code = 4.2, after_UMAP = T, saveddata.dir = saveddata.dir)
```

#### UMAP Visualization
```{r fig.height=4,fig.width=5}
DimPlot(object, reduction = "umap")
```

```{r}
object <- FindNeighbors(object, dims = 1:30)
object <- FindClusters(object, resolution = 0.1)
```

```{r fig.width=12, fig.height=4}
p1 <- DimPlot(object, reduction = "umap", group.by = "geo_accession", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank()) + ggtitle("",subtitle = "GSE154600 (ovary 2)")

p2 <- DimPlot(object, reduction = "umap", group.by = "seurat_clusters", alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())

p3 <- FeaturePlot(object = object, features = c("AIF1"), alpha = 0.5) + NoLegend() + theme(plot.title = element_blank())
p1 + p2 + p3

ggsave("4.2_cell_filtering_UMAP.pdf",plot = p1+p2+p3, device = "pdf", path = paste0(fig.path, "cell_filtering/"), height = 4, width = 12)

```


#### filter Myeloids

```{r}
object_subset <- subset(object, idents = 2)
DimPlot(object_subset, reduction = "umap")
saveRDS(object_subset,file = paste0(saveddata.dir, "Seurat_object_myeloids_integrated/4.2_myeloids_integrated.RDS"))
```



